Misperceived Social Norms about Women Working Outside the Home and Its Effect on Women in Saudi Arabia

### 2024-02-12 outputs/paper/paper.qmd
reference
finish all parts in the paper

---
title: "Misperceived Social Norms about Women Working Outside the Home and Its Effect on Women in Saudi Arabia"
author: 
  - Boxuan Yi
  - Ruoxian Wu
thanks: "Code and data are available at: https://github.com/ScarletWu/Misperceived-Social-Norms-Women-Working-Outside-the-Home-in-Saudi-Arabia. Replication on Social Science Reproduction platform available at: https://www.socialsciencereproduction.org/reproductions/1461/"
date: "February 12, 2024"
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Discussion

## 1. Impact of Informational Intervention:

The research findings have highlighted the importance of correcting misperceived social norms, not only to increase women's participation in the labor market but also to make progress in broader social norms and attitudes towards women's rights and autonomy. The study shows that interventions aimed at challenging and changing societal norms can have a significant impact on women's economic participation and social development. For instance, the study found that after an information intervention, more husbands were willing to sign up their wives for driving lessons, indicating a shift towards more progressive attitudes regarding women's mobility and independence. These changes can contribute to economic development by increasing the labor force and promoting gender equality. The researchers suggest that similar interventions could be effective in other contexts where misperceptions about social norms limit women's economic participation. This approach can offer valuable insights for policymakers and organizations aiming to promote gender equality and economic development, emphasizing the role of accurate information in challenging and changing societal norms. In today's world, digital platforms and social media hold a significant impact on the formation of public opinion and social norms. By making strategic use of these technologies, we can effectively disseminate accurate information and correct any misconceptions. We can leverage online platforms to share data on societal attitudes towards WWOH. This will enable interventions to reach a much wider audience, facilitating a more rapid and widespread change in perceptions.

## 2. Broader Social Proof Implications

It is possible to overcome harmful norms in cultural and societal settings by using social proof.

@@

## Third discussion point

>>>>

---
title: "Misperceived Social Norms about Women Working Outside the Home and Its Effect on Women in Saudi Arabia"
author: 
  - Boxuan Yi
  - Ruoxian Wu
thanks: "Code and data are available at: https://github.com/ScarletWu/Misperceived-Social-Norms-Women-Working-Outside-the-Home-in-Saudi-Arabia. Replication on Social Science Reproduction platform available at: https://www.socialsciencereproduction.org/reproductions/1557/"
date: "February 12, 2024"
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Discussion
## Impact of Informational Intervention: 
The research findings have highlighted the importance of correcting misperceived social norms, not only to increase women's participation in the labor market but also to make progress in broader social norms and attitudes towards women's rights and autonomy. The study shows that interventions aimed at challenging and changing societal norms can have a significant impact on women's economic participation and social development. For instance, the study found that after an information intervention, more husbands were willing to sign up their wives for driving lessons, indicating a shift towards more progressive attitudes regarding women's mobility and independence. These changes can contribute to economic development by increasing the labor force and promoting gender equality. The researchers suggest that similar interventions could be effective in other contexts where misperceptions about social norms limit women's economic participation. This approach can offer valuable insights for policymakers and organizations aiming to promote gender equality and economic development, emphasizing the role of accurate information in challenging and changing societal norms. In today's world, digital platforms and social media hold a significant impact on the formation of public opinion and social norms. By making strategic use of these technologies, we can effectively disseminate accurate information and correct any misconceptions. We can leverage online platforms to share data on societal attitudes towards WWOH. This will enable interventions to reach a much wider audience, facilitating a more rapid and widespread change in perceptions.

## Broader Social Proof Implications
It is possible to overcome harmful norms in cultural and societal settings by using social proof.

@@

## Inverse Pattern of Saudi female labor force participation rate during pandemic
According to the Gender Data Portal from The World Bank, the global average male labor force participation had a decrease of 1.6%, reaching 58.6% in 2020, followed by a rebound to 59.5% in 2021. In fact, this trend was observed across various countries, irrespective of gender.

For example, the Canadian female labor force participation rates in 2019, 2020 and 2021 were 61.4%, 59.5%, 60.6%, respectively, and the Canadian male labor force participation rates for the same years were 70.1%, 68.6%, 69.6%. A similar pattern is discernible in India, where the female labor force participation rates in 2019, 2020 and 2021 were 23.7%, 22.7%, 23%, while corresponding male rates stood at 73.6%, 72.3%, 72.7% [@worldbankLaborForce].

Interestingly, Saudi Arabia exhibited an inverse pattern. While the Saudi male labor force participation rate remained relatively stable, the female rate, along with that of the United Arab Emirates, increased in 2020 and then decreased in 2021. To be more specific, the Saudi female labor force participation rate had a notable increase of 4.9% (nearly one-fifth of the total), reaching 29.5% in 2020, and then declining to 27.8% in 2021. Similarly, female rate in United Arab Emirates increased by 4.2% to 61% in 2020, but decreased to 52.8% in 2021 [@worldbankLaborForce].

The observed fluctuation in world labor force participation rates, particularly Saudi Arabia, leads to speculation on the underlying factors driving these trends. The impact of quarantine is a significant factor, as individuals working from home remained mostly unaffected, unlike those in traditional office settings. This is likely the reason behind the abnormal employment pattern among Saudi women at the start of the pandemic. Moreover, the contrasting pattern in Saudi female labor force participation rate, compared to the world trend, highlights the phenomenon  that many employed females tend to work at home, either willingly or due to traditional social norms. The 2020 surge could indicate potential willingness among Saudi women to explore employment opportunities.
